\phantom{x}\\
\phantom{x}\\
\vspace{1.1em}

\begin{center}
\textbf{\LARGE Acknowledgements}
\end{center} 

First I would like to thank my mother, Judith White, and my siblings David White, Ben White, and Laura White, for their love and support.

I am deeply indebted to the many educators who, during the past twenty one years, have through generosity and patience exposed me to knowledge and taught me how to learn.
First I am thankful to my high school teachers John Donnelly and Eric Dennison for teaching me calculus and computer programming, and Dan Falcon, John Phipps, and Sally Hair for encouraging me to study science.
I want to thank Dartmouth College professors John Thorstensen, Robert Caldwell, and James LaBelle for introducing me to physics and astronomy.
I would also like to thank George O'Toole, Kyle Cady, and John Hammond for involving me with research during this time.

There are many University of Michigan professors that I would like to thank;
Jean Krisch for her guidance;
Dragan Huterer for teaching me introductory physics and supervising me as a freshman;
Sally Oey for teaching me stellar dynamics and supervising me in research as a sophomore;
Christine Aidala for introducing me to high energy experimental physics and supervising me in research as a junior and senior and, along with Leopoldo Pando Zayas, teaching me mathematical methods;
Michael Zieve for teaching me analysis;
Bing Zhou, Roberto Merlin, and James Liu for teaching me the most elegant of theories, electromagnetism;
Bing Zhou again, along with Aaron Leanhardt and Junjie Zhu for teaching me classical mechanics;
Jianming Qian, Tom Schwarz, and Carl Akerlof for teaching me experimental methods;
James Wells, Henriette Elvang, and Ratindranath Akhoury for teaching me quantum field theory.

I have benefited from the help and generosity of many people during my time working on this project.
I would like to thank the members the ATLAS Collaboration who have built and run the detector and computing infrastructure that has enabled my work.
Their work and mine has been supported by various national funding agencies including the Department of Energy.
I would also like to thank the engineers and staff at CERN who help operate the LHC and maintain the site and the ATLAS secretaries who have helped solve numerous problems that occur as a result of living in France.
I am thankful to the Michigan ATLAS group professors Bing Zhou, Jianming Qian, Tom Schwarz, Dan Amidei, Junjie Zhu, Jianming Qian, Homer Neal, and Jay Chapman, as well as Tiesheng Dai, for building a supportive university environment and attracting the funding that has enabled this work.
I am also thankful to faculty of the Laboratoire de l'Acc\^el\^erateur Lin\^eaire and especially Louis and Lydia Fayard.
I am thankful to the friends I have made through physics for engaging in interesting discussions, especially Tom Cheng, Yicheng Guo, Zhi Zheng, Zirui Wang, Rongkun Wang, Zhongyukun Xu, and Siyuan Sun, Yanlin Liu, Aishik Gosh, Aaron White, Ismet Siral, Corentin Allaire, Hao Liu, and Dan Marley.
In particular, Matthew Klein has provided patient and invaluable guidance on topics of data analysis.
During my authorship qualification, I benefited from the guidance of Tom Schwarz and technical instruction of Xueye Hu, Jinhong Wang, and Bob Ball.
On the search for the Higgs decay to two muons, I am thankful for the guidance of Yusheng Wu, Yanlin Liu, and Bing Li.
On the search for contact interactions, I would like to thank the Exotics-LPX Dilepton group and our ever patient editorial board, Paul de Jong, Ulla Blumenschein, and Elliot Lipelles for their feedback.
I would especially like to thank Uta Klein, Dan Hayden, Etienne Dreyer, Peter Falke, Tanya Hryn'Ova, Sean Lawler for their contributions and guidance, and Stephan Hageboeck for his statistical input.
I am thankful for the LPX and Exotics group leadership of Yasu Okumura, Nikolina Llic, Antonio Sidoti, Carl Gwilliam, and Oliver Stelzer-Chilton.
I want to remark on my great fortune to have worked on this project with Deshan Abhayasinghe and Noam Tal Hod, who have made the this project one of the highlights of the past five years.
I am thankful for the committee, Bing Zhou, Tom Schwarz, Christine Aidala, James Wells, and David Baker for reviewing this thesis.
Finally, I want to extend my deepest thanks to my advisor, Bing Zhou.
She has worked tirelessly to support me since, as an undergraduate, I enrolled in her section of Physics 405: Intermediate Electromagnetism.
As a role model, her persistent curiosity and attention to detail have provided an exceptional example to follow of an experimental physicist. 
As a mentor, she has always pushed me to work to my fullest ability.


